---
title: "Space-time prism trajectories"
author: "Mark ten Vregelaar"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 5


vignette: >
  %\VignetteIndexEntry{Space-time prism trajectories}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction


This vignette is meant to show an example of what is possible with the STPtrajectories package. In this vignette two STP_Tracks are created and Analysed. Altough the tracks are randomly generated, they are meant to reperesent two bears. 
The following steps will be implemeted:

* Create two pairs of space-time points using the STP_Track class. 
* Add more points by applying the random trajectory generator(RTG)
* Calculate the PPA area for both tracks
* Apply the alibi query
* Visulise the STPtrajectories


Most functions avaliable in the package will be used in this example.

```{r,include=FALSE}
library(STPtrajectories)
```
<center>__The functions of the STPtrajectories package:__</center>


| Functions     |               |
| ------------- |:-------------:|
| alibi_query   | STP_plot      |
| axes_STP_plot | PPA           |
| RTG           | STP_Track     |
| getVmaxtrack  | potential_stay|

# The example project


First Load the required packages
```{r setup}
library(sp)
library(spacetime)
library(knitr)
library(rgl)
library(rgeos)
knit_hooks$set(webgl = hook_webgl)
```


## Create two pairs of space-time points
```{r}
crs_NL = CRS("+init=epsg:28992")
t1 <- as.POSIXct(strptime("01/01/2017 12:00:00", "%m/%d/%Y %H:%M:%S"))
t2 <- as.POSIXct(strptime("01/01/2017 15:00:00", "%m/%d/%Y %H:%M:%S"))
time<-c(t1,t2)


# Spatial coordinates
x_bear1=c(7500,10);y_bear1=c(10,6000)

x_bear2=c(20,8000);y_bear2=c(30,7000)

n = length(x_bear1)

# create class STIDF {spacetime}
stidf_bear1 = STIDF(SpatialPoints(cbind(x_bear1,y_bear1),crs_NL), time, data.frame(co2 = rnorm(n),O2=rnorm(n)))
stidf_bear2 = STIDF(SpatialPoints(cbind(x_bear2,y_bear2),crs_NL), time, data.frame(co2 = rnorm(n),O2=rnorm(n)))

# Track-class {trajectories}
track_bear1<-Track(stidf_bear1)
track_bear2<-Track(stidf_bear2)
# Set maximum speed 
# max speed for a period of 3 hours in m/s
v_bear1<-5/3.6
v_bear2<-6/3.6
# STP_track class
STP_bear1<-STP_Track(track_bear1,v_bear1)
STP_bear2<-STP_Track(track_bear2,v_bear2)
# plot the points
plot(STP_bear1,type='p',col='red',pch=16,cex=2,xlim=c(0,9000),ylim=c(0,8000),xlab='x (meters)',
     ylab='y (meters)', main= 'Start and end locations of bear trajecories')
plot(STP_bear2,type='p',col='blue',pch=16,cex=2,add=T)
legend('topright',c('Bear 1','Bear 2'),pch = 16,col=c('red','blue'))
text(c(x_bear1,x_bear2), c(y_bear1,y_bear2), labels=c('start','end'), cex= 1.2,adj= c(0.5,-1))
```
 
## Apply random trajectory generator(RTG)
```{r}
## Create trajecotires using the RTG
# set seed to create same trajecory
set.seed(10)
STP_track_bear1<-RTG(STP_bear1,n_points = 15)
set.seed(2)
STP_track_bear2<-RTG(STP_bear2,n_points = 15)

# plot results
plot(STP_track_bear1,type='b',col='red',pch=16,cex=0.8,xlim=c(-1000,12000),ylim=c(-500,9000),xlab='x (meters)',ylab='y (meters)', main= 'The two bear trajecories')
plot(STP_track_bear2,type='b',col='blue',pch=16,cex=0.8,add=T)
legend('topright',c('Bear 1','Bear 2'),pch = 16,lty =1, col=c('red','blue'))

```

## Calculate the Potential Path Area 
The time difference between the space-time points is now a bit over 11 minutes and thus are the maximum speeds of 5 and 6 km/h not realsitic. A more realstic max speed is 1.5 times the speed required to reach every point.
```{r}
# set the new maximum speed. Same for every segment
vmax_bear1<-getVmaxtrack(STP_track_bear1)*1.5
vmax_bear2<-getVmaxtrack(STP_track_bear2)*1.5

STP_track_bear1@connections$vmax<-vmax_bear1
STP_track_bear2@connections$vmax<-vmax_bear2

# calculate Potential Path Area (PPA)
PPA_bear1 <- PPA(STP_track_bear1)
PPA_bear2 <- PPA(STP_track_bear2)


# plot results
plot(STP_track_bear1,type='b',col='red',pch=16,cex=0.8,xlim=c(-1000,12000),ylim=c(-500,9000),xlab='x (meters)',ylab='y (meters)', main= 'The two bear trajecories with PPA')
plot(STP_track_bear2,type='b',col='blue',pch=16,cex=0.8,add=T)
legend('topright',c('Bear 1','Bear 2'),pch = 16,lty =1, col=c('red','blue'))


plot(PPA_bear1,add=T)
plot(PPA_bear2,add=T)

```
## the alibi query

The figure above shows a spatial intersetion between the trajectories. To test wether they could have been at the same location at the same time you can apply the alibi query.
```{r alibi_query}
alibi_query(STP_track_bear1,STP_track_bear2,stop_if_true = T)# not always correct. package still in development

```
## Visulise STP_tracks
```{r STP_plot, webgl=TRUE}
zfac<- 50 # aspect ration between sptatial axes and time axis
t_int <- 0.8 # determines how many PPAs are used to visualise STPs.
open3d()
STP_plot(STP_track_bear1,time_interval = t_int,zfactor = zfac)
STP_plot(STP_track_bear2,time_interval = t_int,zfactor = zfac,st = STP_track_bear1@endTime[1],col = 'blue')

# add axes
#axes_STP_plot(time,z_factor = zfac) function to add axes but not suitable for Rmarkdown
title3d(main = "3D Visualisation of STP_tracks",xlab='x',ylab='y',cex=1.3)
bg3d('lightblue')
# data time axis
tdif<-as.numeric(difftime(time[2],time[1],units = 'mins'))
tickval<-seq(0,tdif*zfac,length.out = 5)
timesval<-seq(time[1],time[2],length.out = 5)
# add axes
axes3d(c('x','y'),xlab='x')
axis3d('z',at=tickval,labels = timesval,cex=0.8)
box3d()
```

## potential stay time
Somone saw a bear on a road somewhere between 12 and 1 o'clock. The code below shows how you can 
check if bear 1 was that bear and when he could have been at the road

```{r, webgl=TRUE}
bear1_sub <-STP_track_bear1[1:20,'2017-01-01 12:00:00 CET::2017-01-01 13:00:00 CET']

road<-readWKT("LINESTRING(2800 500,3200 200,4000 50,5000 100,6000 500,7000 1000,7700 1200)")

road@proj4string<-crs_NL
road_buffer<-gBuffer(road,width = 4)

intervals<-potential_stay(STP_track_bear1,road_buffer)
intervals

# calculate the time the individual could have been at the lake
road_time <- sum(sapply(intervals, function(int){difftime(int[2],int[1],units = 'mins')}))

print(paste('Total time bear 1 could have been on the road is ',round(road_time,2),'minutes'))


# visulise in 3D
open3d()
zfac<-STP_plot(bear1_sub,time_interval = 0.6)

axes_STP_plot(c(bear1_sub@endTime[1],tail(bear1_sub@endTime,1)),z_factor = zfac,n_ticks_z = 5,n_ticks_xy = 4)


x<-road_buffer@polygons[[1]]@Polygons[[1]]@coords[,1]
y<-road_buffer@polygons[[1]]@Polygons[[1]]@coords[,2]
z<-difftime(tail(bear1_sub@endTime,1),bear1_sub@endTime[1],units = 'mins')*zfac


shade3d(translate3d(
  extrude3d(x,y,thickness = z),0,0,0),col='black',add=TRUE)


```


## create rough space-time prisms
```{r, webgl=TRUE}
# higher maximum speed 
vmax<- bear1_sub@connections$vmax[1]*1.6
# taking into account uncerainty about location and measurement time at control points.
open3d()
bear1_sub_rough<-STP_Track(bear1_sub,vmax,0,location_uncertainty = 2,time_uncertainty = 1)

zf<-STP_plot(bear1_sub_rough,alpha = 0.4,col='darkcyan')
STP_plot(bear1_sub,col='green',zfactor = zf,st=bear1_sub_rough@endTime[1]-1*60)
axes_STP_plot(c(bear1_sub_rough@endTime[1]-1*60,bear1_sub_rough@endTime[6]+1*60),z_factor = zf)

```

See also:

[github](https://github.com/markvregel/STPtrajectories)
